\documentclass[11pt]{article}

\include{init}

\title{Bayesian Approach to Inverse Problems: an Application to NNPDF Closure Testing}
\author[a]{Luigi Del Debbio} 
\author[b]{Tommaso Giani} 
\author[a]{Michael Wilson}
\affil[a]{Higgs Centre for Theoretical Physics, School of Physics and Astronomy,
Peter~Guthrie~Tait~Road, Edinburgh EH9 3 FD, United Kingdom.}
\affil[b]{Nikhef Theory Group, Science Park 105, 1098 XG Amsterdam, The Netherlands}


\makeindex

\begin{document}

\maketitle

\begin{abstract}
    We discuss the Bayesian approach to the solution of inverse problems and
    apply the formalism to analyse the closure tests performed by the NNPDF
    collaboration. Starting from a comparison with the approach that is
    currently used for the determination of parton distributions (PDFs) by the
    NNPDF collaboration, we discuss some analytical results that can be obtained
    for linear problems and use these results as a guidance for the more
    complicated non-linear problems. We show that, in the case of Gaussian
    distributions, the posterior probability density of the parametrized PDFs is
    fully determined by the results of the NNPDF fitting procedure. In the
    particular case that we consider, the fitting procedure and the Bayesian
    analysis yield exactly the same result. Building on the insight that we
    obtain from the analytical results, we introduce new estimators to assess
    the statistical faithfulness of the fit results in closure tests. These
    estimators are defined in data space, and can be studied analytically using
    the Bayesian formalism in a linear model in order to clarify their meaning.
    Finally we present numerical results from a number of closure tests
    performed with current NNPDF methodologies. These further tests allow us to
    validate the new methodology and provide a quantitative comparison of the
    new and old methodologies. As PDFs determinations move into precision
    territory, the need for a careful validation of the methodology becomes
    increasingly important: the error bar has become the focal point of
    contemporary PDFs determinations. In this perspective, theoretical
    assumptions and other sources of error are best formulated and analysed in
    the Bayesian framework, which provides an ideal language to address the
    precision and the accuracy of current fits. 

\end{abstract}

\include{introduction}
\include{inverse}
\include{closure_test}
\include{estimators}
%\include{models}
%\include{experimental_setup}
\include{results}
\include{summary}
\appendix
\include{gaussian}
\include{app_datasets}
\include{app_deltachi2}
\bibliographystyle{unsrt}
\bibliography{biblio}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
